Promise and peril: Agnostic identification methods\\ for detecting differential item functioning

### 2020-03-14 paper/paper.tex
after knitting and sending that copy to ben d and just before doing a reorganization to try to keep this repo wieldy

\begin{document}
\maketitle
\begin{abstract}
It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models. However, to use an LRT, the model needs to be identified so that differences in group ability can be disentangled from potential DIF. We summarize existing assumptionless identification (AI) methods and propose a variety of new methods. We conduct a simulation study --- which we believe to be more realistic than most DIF simulation studies in the literature --- and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods. We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that presents clearly all information about possible DIF, including sampling variability in item parameters, to the analyst. \clearpage
\end{abstract}

@@

We find it informative to describe simulation conditions in terms of multidimensional abilities as opposed to item parameters that vary across groups. For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously \({b_j}^{\text{ref}}\) for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for all students in the reference group and \({\eta_i}^{\text{ref}} = -1\) for all students in the focal group. We now need to use a compensatory 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional compensatory 2PL model, the probability student \(i\) responding correctly to item \(j\) is
\[
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
\]
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability across students in the same group.

@@

\hypertarget{assumptionless-identification-methods}{%
\section{Assumptionless identification methods}\label{assumptionless-identification-methods}}

Psychometricians have long been in search of the perfect DIF detection method. It has become well known that the IRT-based likelihood ratio test (LRT) is the best way to test for DIF, which has been well demonstrated (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}). As a result, we avoid methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which muddies the waters by moving away from the IRT framework, and has been shown to perform no better --- and in some cases worse --- than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990}).

The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions --- about either relative group ability or which items contain bias --- can be made. That is, we're interested in assumptionless identification methods (hereafter referred to as \enquote{AI methods}) so that the LRTs can be used to test items for DIF. For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection. And, in search of a coherent framework, we sometimes edit names of existing methods. We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to the proliferation of names.

In this section, we summarize existing AI methods and propose extensions of those methods. We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test. Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{ref}} = -1\). The slope on target ability is set to \(a_{1j} = 1\) for all items. The slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise. From the DIF-as-item-property view, there is no nuisance ability. Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.

In general, we denote the mean ability of the reference and focal group as \(\mu^\text{ref}\) and \(\mu^\text{foc}\), respectively. We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\). The fundamental challenge is to use an AI method to identify the model, which allows for the disentangling \(\hat\mu^\text{foc}\) from estimating \(\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\). The most common AI method is to select a group of anchor items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT.

Computing is done in R (R Core Team \protect\hyperlink{ref-rcore}{2019}), model fitting in the R package mirt (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}), and data wrangling/visualization in the set of R packages known as the tidyverse (Wickham \protect\hyperlink{ref-tidy}{2017}). Code is available at TO DO.

@@

\hypertarget{all-others-as-anchors}{%
\subsubsection{All-others-as-anchors}\label{all-others-as-anchors}}

Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used AI methods and unequivocally recommended the all-others-as-anchors (AOAA) method. AOAA tests each item for DIF one at a time using all of the other items as anchors. For example, when testing the first item for DIF, all of the other items are used as anchors. This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}). Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on. The items for which the flexible model outperforms the baseline model (typically based on a \(\chi^2\) test) are identified as having DIF, and the rest of the items become anchor items. Note that AOAA is implemented in the mirt R package, and is called by passing scheme = \enquote{drop} to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).

@@

\hypertarget{all-others-as-anchors-all-significant}{%
\paragraph{All-others-as-anchors-all-significant}\label{all-others-as-anchors-all-significant}}

One way to attempt to counter artificial DIF is with purification through iterative anchor selection. For example, Drasgow (\protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively --- with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model --- until no more items tested positively. Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free. We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS). Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set. Note that AOAA-AS is implemented in the mirt R package, and is called by passing scheme = \enquote{drop\_sequential} to the DIF function.

AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF? With a sufficient sample size and at least one item with DIF, this will necessarily be the case. In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items. Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly. Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF. It's unclear if one should proceed if those four items display DIF.

\hypertarget{all-other-as-anchors-one-at-a-time}{%
\paragraph{All-other-as-anchors-one-at-a-time}\label{all-other-as-anchors-one-at-a-time}}

@@

Applying AOAA-OAT to our thought experiment demonstrates its effectiveness. The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF. In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion. AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.

@@

Table \ref{table:themethods} summarizes the three all-others-as-anchors methods. It's useful to remember that AOAA is not an iterative procedure. The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method. In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified. On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.

\begin{table}[H]
\caption{Summary of the three all-others-as-anchors assumptionless identification methods}
\centering
\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
\toprule

@@

\hypertarget{clustering-items}{%
\subsubsection{Clustering items}\label{clustering-items}}

@@

\hypertarget{the-equal-means-multiple-imputation-logit-graph}{%
\subsubsection{The equal means, multiple imputation logit graph}\label{the-equal-means-multiple-imputation-logit-graph}}

@@

We propose a new method, the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst. Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting \(\mu^\text{foc}\) to \(0\). Again, the result is that all differences in performance manifest in \(\tilde{\mathbf{d}}\). To measure the variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}). Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of \(\tilde{d_j}\) for each item. These are the distributions displayed in a EM-MILG. The method is inspired in part by Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.

Figure \ref{fig:emmilg} shows the EM-MILG for our simulated data. We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance. The challenge, then, is to select the anchor items. The analyst might assume that --- because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and can be used as anchor items.

@@

After choosing the anchor items, the model is refit. The new model is identified by setting \(d_j = 0\) for the anchor items, instead of by setting \(\mu^\text{foc}\) to \(0\). The same process of using multiple imputations to estimate the distribution of \(d_j\) can be used with the new model. Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG). As expected, in our one-run simulation, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.

@@

One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security. Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items. A EM-MILG combats this concern by presenting all information clearly to the analyst. In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased. Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data. And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG. In particular, we could visualize DIF in the non-anchor items after using AOAA or CLUSTER to identify the model.

@@

The previously discussed AI methods select a set of anchor items, whether it is an algorithm or the analyst that makes that selection. The anchor items are used to estimate \(\hat \mu^\text{foc}\). An alternative strategy is to directly select the anchor point, \(\mu^{\star\text{foc}}\). Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, all items can be tested for DIF. The question then becomes the following: How is the anchor point selected?

\hypertarget{gini-index}{%
\subsubsection{Gini index}\label{gini-index}}

@@

\(\mu^{\star\text{foc}}\) is selected by maximizing the Gini index (which is why we use the abbreviation MGI). The intuition and assumption is that some items might have bias but the majority do not. Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), MGI sets
\[
\mu^{\star\text{foc}} = \mathop\mathrm{arg\,max}_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
\]
where \(\mu^\text{foc} \in (-\infty, \infty)\) and \(\mu^\text{foc}\) is added to each element of \(\mathbf{d}\).

For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible \(\mu^\text{foc}\) values. The result of MGI is \(\mu^{\star\text{foc}} = -0.99\).

@@

\caption{Maximizing the Gini index (MGI) to select the anchor point.}\label{fig:ginipath}
\end{figure}

The model is then refit with the model identifying assumption of \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\), and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.

@@

\caption{The MILG with $\mu^{\text{foc}}$ set to $\mu^{\star\text{foc}} = -0.99$}\label{fig:ginimilg}
\end{figure}

\hypertarget{minimizing-between-curves}{%
\subsubsection{Minimizing between curves}\label{minimizing-between-curves}}

Raju's area method (Raju \protect\hyperlink{ref-raju1988area}{1988}) measures the amount of DIF by calculating the area between the item characteristic curves, the function that maps the student's ability to their probability of correct response, of the two groups:
\[
\text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
\]
Raju's area method, along with AOAA, has been cited as the two most commonly used IRT-based DIF detection methods (Magis et al. \protect\hyperlink{ref-magis2011generalized}{2011}). We see two weakness of Raju's area method: First, the item characteristic curves still need to be linked by anchor items or an anchor point, which the method does not help to identify. Second, the area is unweighted, so all values of \(\theta\) matter equally, despite some being much more realistic than others.

To build off of Raju's area method, we propose a new method, which we call \enquote{minimizing the area between curves} (MABC). To understand MABC, imagine multigroup item response data where \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\) so that there is no DIF. The fundamental identification problem of DIF in multigroup IRT models is that there are an infinite number of models with the same likelihood from which to choose. For example, we could correctly assume that the focal group has the same ability as the reference group and fix \(\mu^{\star\text{foc}} = 0\). The model would then estimate \(\hat d_j \approx 0 \forall j\), and we would correctly conclude the groups have the same ability and there is no DIF. Alternatively, we could assume that the focal group has \(\mu^{\star\text{foc}} = 3\). The model would then compensate by finding \(\hat d_j \approx -3 \forall j\), and we would incorrectly conclude that the focal group is high ability, but every item contains DIF against them. Both of these models have the same likelihood, so how should one choose which model to believe? MABC chooses the model with the least amount of total DIF, as measured by the total weighted area between the item characteristic curves. As a result, the likelihood tie is broken by preferring to explain differences in performance across groups by ability differences as opposed to DIF.

>>>>

\begin{document}
\maketitle
\begin{abstract}
It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models. However, to use an LRT, the model needs to be identified so that differences in group ability can be disentangled from potential DIF. We summarize existing agnostic identification (AI) methods and propose a variety of new methods. We conduct a simulation study --- which we believe to be more realistic than most DIF simulation studies in the literature --- and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods. We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that presents clearly all information about possible DIF, including sampling variability in item parameters, to the analyst. \clearpage
\end{abstract}

@@

For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional. For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously \({b_j}^{\text{ref}}\) for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for all students in the reference group and \({\eta_i}^{\text{ref}} = -1\) for all students in the focal group. We now need to use a compensatory 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional compensatory 2PL model, the probability student \(i\) responds correctly to item \(j\) is
\[
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
\]
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.

@@

\hypertarget{agnostic-identification-methods}{%
\section{Agnostic identification methods}\label{agnostic-identification-methods}}

Psychometricians have long been in search of the perfect DIF detection method. It has been well demonstrated that the IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}). As a result, we avoid methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which muddies the waters by moving away from the IRT framework, and has been shown to perform no better --- and in some cases worse --- than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990}).

The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions --- about either relative group ability or which items might have DIF --- can be made. That is, we're interested in agnostic identification methods (hereafter referred to as \enquote{AI methods}) so that LRTs can be used to test items for DIF. For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection. And, in search of a coherent framework, we sometimes edit names of existing methods. We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.

In this section, we summarize existing AI methods and propose extensions of those methods. We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test. Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{foc}} = -1\). The slope on target ability is set to \(a_{1j} = 1\) for all items. The slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise. We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability. Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.

In general, we denote the mean ability of the reference and focal group as \(\mu^\text{ref}\) and \(\mu^\text{foc}\), respectively. We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\). The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of \(\hat\mu^\text{foc}\) from the estimation of \(\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\). The most common AI approach is to select a group of anchor items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT.

Computing is done in R (R Core Team \protect\hyperlink{ref-rcore}{2019}), model fitting in the mirt R package (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}), and data wrangling/visualization in the suite of R packages known as the tidyverse (Wickham \protect\hyperlink{ref-tidy}{2017}). Code is available at \emph{(TODO)}.

@@

\hypertarget{all-others-as-anchors-aoaa}{%
\subsubsection{All-others-as-anchors (AOAA)}\label{all-others-as-anchors-aoaa}}

Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used AI methods and unequivocally recommended the all-others-as-anchors (AOAA) method. AOAA tests each item for DIF one at a time using all of the other items as anchors. For example, when testing the first item for DIF, all of the other items are used as anchors. This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}). Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on. The items for which the flexible model outperforms the baseline model (typically based on a \(\chi^2\) test) are identified as having DIF, and the rest of the items become anchor items. AOAA is implemented in the mirt R package, and is called by passing scheme = \enquote{drop} to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).

@@

\hypertarget{all-others-as-anchors-all-significant-aoaa-as}{%
\paragraph{All-others-as-anchors-all-significant (AOAA-AS)}\label{all-others-as-anchors-all-significant-aoaa-as}}

One way to attempt to counter artificial DIF is with purification through iterative anchor selection. For example, Drasgow (\protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively --- with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model --- until no more items tested positively. Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free. We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS). Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set. AOAA-AS is implemented in the mirt R package, and is called by passing scheme = \enquote{drop\_sequential} to the DIF function.

AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF? With a sufficient sample size and at least one item with DIF, this will necessarily be the case. In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items. Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly. Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF. It's unclear if one should proceed if those four items test positive for DIF, too.

\hypertarget{all-other-as-anchors-one-at-a-time-aoaa-oat}{%
\paragraph{All-other-as-anchors-one-at-a-time (AOAA-OAT)}\label{all-other-as-anchors-one-at-a-time-aoaa-oat}}

@@

Applying AOAA-OAT to our thought experiment demonstrates its effectiveness. The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF. In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion. To work, AOAA-OAT has two requirements: First, that at least two items do not have DIF, and second, that the set of items without DIF are more homogeneous than other sets of items.

@@

Table \ref{table:themethods} summarizes the three all-others-as-anchors methods. It's useful to remember that AOAA is not an iterative procedure. The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method. In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified. On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.

\begin{table}[H]
\caption{Summary of the three all-others-as-anchors agnostic identification methods}
\centering
\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
\toprule

@@

\hypertarget{equal-means-clustering-emc}{%
\subsubsection{Equal means clustering (EMC)}\label{equal-means-clustering-emc}}

@@

\hypertarget{the-equal-means-multiple-imputation-logit-graph-em-milg}{%
\subsubsection{The equal means, multiple imputation logit graph (EM-MILG)}\label{the-equal-means-multiple-imputation-logit-graph-em-milg}}

@@

We propose a new method, the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst. Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting \(\mu^\text{foc}\) to \(0\). Again, the result is that all differences in performance manifest in \(\tilde{\mathbf{d}}\). To measure the variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}). Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are drawn to estimate the distribution of \(\tilde{d_j}\) for each item. These are the distributions displayed in a EM-MILG. The method is inspired in part by Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.

Figure \ref{fig:emmilg} shows the EM-MILG for our one-run simulation. We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance. The challenge, then, is to select the anchor items. The analyst might assume that --- because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and can be used as anchor items.

@@

After choosing the anchor items, the model is refit. The new model is identified by setting \(d_j = 0\) for the anchor items, instead of by setting \(\mu^\text{foc} = 0\). The same process of using multiple imputations to estimate the distribution of \(\tilde{d_j}\) can be used with the new model. Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG). As expected, in our one-run simulation, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.

@@

One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security. Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items. A EM-MILG combats this concern by presenting all information clearly to the analyst. In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased. Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data. And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG. In particular, we could visualize DIF in the non-anchor items after using, for example, AOAA-OAT or EMC to identify the model.

@@

The previously discussed AI methods select a set of anchor items, whether it is an algorithm or the analyst that makes that selection. The anchor items are used to estimate \(\hat \mu^\text{foc}\). An alternative strategy is to directly set the anchor point, \(\mu^{\star\text{foc}}\). Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF. The question then becomes the following: How is the anchor point selected?

\hypertarget{maximizing-the-gini-index-maxgi}{%
\subsubsection{Maximizing the Gini index (MAXGI)}\label{maximizing-the-gini-index-maxgi}}

@@

\(\mu^{\star\text{foc}}\) is selected by maximizing the Gini index (thus the abbreviation MAXGI). The intuition and assumption is that the items without DIF are the most homogeneous. Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), MGI sets
\[
\mu^{\star\text{foc}} = \mathop\mathrm{arg\,max}_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
\]
where \(\mu^\text{foc} \in (-\infty, \infty)\) and \(\mu^\text{foc}\) is added to each element of \(\mathbf{d}\).

For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible \(\mu^\text{foc}\) values. The result of MAXGI is \(\mu^{\star\text{foc}} = -0.99\), which is extraordinarily close to the data-generating value of -1.

@@

\caption{Maximizing the Gini index (MAXGI) to select the anchor point.}\label{fig:ginipath}
\end{figure}

The model is then refit with the identifying assumption that \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\), and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.

@@

\caption{The MILG with $\mu^{\text{foc}}$ set to $\mu^{\star\text{foc}} = -0.99$.}\label{fig:ginimilg}
\end{figure}

\hypertarget{minimizing-between-curves-minbc}{%
\subsubsection{Minimizing between curves (MINBC)}\label{minimizing-between-curves-minbc}}

Raju's area method (Raju \protect\hyperlink{ref-raju1988area}{1988}) measures the amount of DIF by calculating the area between the item characteristic curves, the function that maps the student's ability to their probability of correct response, of the two groups:
\[
\text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
\]
Raju's area method has been cited as one of the most commonly used IRT-based DIF detection methods (Magis et al. \protect\hyperlink{ref-magis2011generalized}{2011}). However, Raju's area method is not an AI method because the item characteristic curves still need to be linked by anchor items or an anchor point. An additional weakness is that the area is unweighted, so all values of \(\theta\) matter equally, despite some being much more realistic than others.

To adapt Raju's area method into an AI method, we propose a new method, which we call \enquote{minimizing the area between curves} (MINBC). To understand MINBC, imagine a scenario in which the data-generating process is \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\), so that there is no DIF. The fundamental identification problem is that there are an infinite number of models with the same likelihood from which to choose. For example, we could correctly assume that the focal group has the same ability as the reference group and fix \(\mu^{\star\text{foc}} = 0\). The model would then estimate \(\hat d_j \approx 0 \forall j\), and we would correctly conclude the groups have the same ability and there is no DIF. Alternatively, we could assume that the focal group has \(\mu^{\star\text{foc}} = 3\). The model would then compensate by finding \(\hat d_j \approx -3 \forall j\), and we would incorrectly conclude that the focal group is high ability, but every item contains DIF against them. Both of these models have the same likelihood, so how should one choose which model to believe? MINBC chooses the model with the least amount of total DIF, as measured by the total weighted area between the item characteristic curves. As a result, the likelihood tie is broken by preferring to explain differences in performance across groups by ability differences (as opposed to DIF).

Denote a function that takes \(\mu^\text{foc}\) as input and estimates \(\hat b_j^\text{foc}\) by fitting a unidimensional Rasch model as \(m_j(\mu^\text{foc})\). The amount of DIF on each item is calculated as
\[
\text{DIF}_j(\mu^\text{foc}) = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, m_j(\mu^\text{foc}))| g(\theta)d\theta
\]
where \(g(\theta)\) is a weighting function such that \(\int g(\theta)d\theta = 1\). The total DIF on the test, then, is
\[
\text{Total DIF}(\mu^\text{foc}) = \sum_{j} \text{DIF}_j(\mu^\text{foc})
\]
In this way, \(\text{Total DIF}(\mu^\text{foc})\) is a function where the input is \(\mu^\text{foc}\) and the output is the total amount of DIF on the test. MINBC sets
\[
\mu^{\star\text{foc}} = \mathop\mathrm{arg\,min}_{\mu^\text{foc}} \text{Total DIF}(\mu^\text{foc}).
\]
MINBC is inspired in part by Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}), who use the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF). The selection of \(g(\theta)\) results in the relative weighting of \(\theta\) values. Chalmers, Counsell, and Flora do not discuss how to choose \(g(\theta)\) and in their empirical examples use \(g(\theta)\) uniform for \(-6 \le \theta \le 6\), which may be suboptimal in some cases. It might seem intuitive to choose \(g(\theta) \sim N(0, 1)\) because \(\mu^\text{ref} = 0\), but this choice doesn't take into account the ability distribution of the focal group. If \(\mu^\text{foc} = 3\), wouldn't we also want to prioritize high \(\theta\) values? Accordingly, we set \(g(\theta)\) to be the average of the reference and focal group ability probability density functions:
\[
g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
\]
